---
title: "Working with bin files"
author: "Caliper Staff"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with bin files}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(caliperR)
dk <- connect()
```

## Intro

Caliper software makes use of a number of custom file formats. Tabular data is
routinely stored in fixed-format binary (FFB) files with extension ".bin". The
`caliperR` package allows you to read and write these files even if you don't
have Caliper software installed on your machine (although it is faster if you
do).

## Reading bin data

The `caliperR` package comes with a sample bin file for this vignette.

```{r}
bin_file <- system.file(
  "extdata", "gisdk", "testing", "toy_table.bin", package = "caliperR"
)
```

Use the `read_bin()` function to bring it into R.

```{r}
df <- read_bin(bin_file)
df
```

### Field descriptions

If the bin file has field descriptions, they are preserved. They will appear
when using `View()` in Rstudio, but can also be seen using the Hmisc package.

```{r}
# View(df)

descriptions <- Hmisc::label(df)
descriptions
```

### Field display names

Lastly, the `caliperR` package can also read and return any display names that
may be assigned to the FFB table.

```{r}
display_names <- read_bin(bin_file, returnDnames = TRUE)
display_names
```

You can substitute display names for the field names using `colnames()`.

```{r}
df2 <- df
colnames(df2) <- display_names
df2
```

## Writing bin data

Use the `write_bin()` function to write the data back out to a bin file. The
`descriptions` and `display_names` arguments are optional.

```{r}
temp_bin <- tempfile(fileext = ".bin")
write_bin(df, temp_bin, descriptions, display_names)
```

