---
title: "Caliper matrices"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Caliper matrices}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(caliperR)
dk <- connect()
```

## Intro

This vignette explains how to work with Caliper matrices from R using the
`caliperR` package. Many of the relevant GISDK functions have been streamlined
using the `CaliperMatrix` R6 class and concepts such as matrix "handles" and
"currencies" are resolved behind the scenes. The end result is a more intuitive
experience.

This vignette assumes you are already familiar with the basics of the `caliperR`
package. If not, see `vignette("using-caliperR")`.

Before getting started, I create a temporary copy of this package's matrix file. 
This is only to ensure that changes made by this vignette are not made to the
original file (and is not necessary in normal workflow).

```{r}
orig <- system.file("extdata", "gisdk", "testing", "toy_matrix.mtx", package = "caliperR")
mtx_file <- tempfile(fileext = ".mtx")
ok <- file.copy(orig, mtx_file)
```

## Getting started

Opening a matrix returns a `CaliperMatrix` object with useful attributes and
methods.

```{r}
mtx <- dk$OpenMatrix(mtx_file, NA)
```

Several of R's generic functions can be used on the resulting object.

```{r}
summary(mtx)
```

```{r}
head(as.data.frame(mtx))
```

The function `as.matrix()` can be used on a specific core, or on the matrix object
directly to convert all cores.

```{r}
as.matrix(mtx$core_b)
as.matrix(mtx)
```

You can also send updated data back to the Caliper matrix. In the code below,
I update the top left corner of `core_b`. Doing so updates the .mtx file.

```{r}
new_data <- matrix(seq(1, 4), nrow = 2, ncol = 2, 
                   dimnames = list(c(1,2), c(1,2)))
new_data
mtx$core_b <- new_data
as.matrix(mtx$core_b)
```

### Full work flow example

The example below introduces one more function: `CreateCore()`. Using it, we
create a new core (`core_c`) and fill it with the sum of `core_a` and `core_b`.

```{r}
mtx$CreateCore("core_c")
data <- as.matrix(mtx)
data$core_c <- data$core_a + data$core_b
mtx$core_c <- data$core_c
as.matrix(mtx$core_c)
```

## Working with indices

Row and column indices are how you subset Caliper matrices. In the previous
section, we were able to update a subset of matrix cells directly
from R without needing to know about matrix indices; however, knowledge of
indices is valuable when working with more complex GISDK functions.

You can list the indices availabe in the matrix:

```{r}
mtx$indices
```

The "all" index contains all rows and columns 1-5. The "subset" index is only
1-3.

The `row_index` and `column_index` active fields allow you to see which index is
active or switch between available indices. First, check the currently-active
row index:

```{r}
mtx$row_index
```

The "all" row index is active, which means all 5 rows. Change it to the "subset"
index to only work with the first 3.

```{r}
mtx$row_index <- "subset"
as.matrix(mtx$core_b)
```

We can do the same to the column index, which gives us a 3x3 matrix.

```{r}
mtx$column_index <- "subset"
as.matrix(mtx$core_b)
```

Finally, the `caliperR` package makes it simple to add new indices. The example
below creates an index named "new" and gives the IDs of the rows/columns to
include.

```{r}
mtx$CreateIndex("new", c(1, 2, 3, 4))
mtx$indices
```

## Handles and currencies

Note that none of the examples above required that you know what a matrix handle
or currency is The `caliperR` package handles this behind the scenes, but some
GISDK functions require handles or currencies as arguments.

If a function requires the matrix handle, simply pass in the entire matrix
object. `caliperR` will understand what you mean.

```{r}
# A GISDK function requiring a matrix handle
dk$GetMatrixBaseIndex(mtx)
```

The function `GetMatrixColumnLabels()` requires a matrix currency. Simply
reference the core by name and `caliperR` handles the rest. Note that the
column labels are only 1-3 as opposed to 1-5. The `subset` column index we set
is respected.

```{r}
# A GISDK function requiring a matrix currency
dk$GetMatrixColumnLabels(mtx$core_a)
```
