---
title: "Caliper objects"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Caliper objects}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(caliperR)
dk <- connect()
```

## Intro

The Caliper scripting language (GISDK) contains numerous objects which
facilitate the analysis of transport problems. This vignette explains how to
work with these objects from R using the `caliperR` package. If you are not
already familiar with the basics of this package, see
`vignette("using-caliperR")`.

## Getting started

In TransCAD, the `NLM.Model` object facilitates the creation and application of
logistic regression models. To use this object from R, we create it like so:

```{r}
object <- CreateObject("NLM.Model")
```

This creates both the GISDK object and an R object (R6 class) that points to it.
From here, the only difference between R and GISDK is syntax: R uses `$` instead
of `.` to access object attributes and methods. The code below sets the `Label`
attribute of the `NLM.Model` GISDK object.

```{r}
object$Label <- "A logit model"
object$Label
```

Next, the `Clear()` method of the `NLM.Model` object is called, which removes
the value of all fields including `Label`.

```{r}
object$Clear()
object$Label
```

To see the complete list of available fields and methods for the GISDK object,
simply use `object$info`.

```{r}
info <- object$info
head(info$FieldNames)
head(info$MethodNames)
```

