---
title: "Using caliperR"
author: "Caliper Staff"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Using caliperR}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

`caliperR` makes it easy to work with Caliper software from R. Supported
software includes:

  * [TransCAD](https://www.caliper.com/tcovu.htm)
  * [TransModeler](https://www.caliper.com/transmodeler/default.htm)
  * [Maptitude](https://www.caliper.com/maptovu.htm)

The sections below show how to work with the package.

## Connecting

Connecting to Caliper software is accomplished by calling the `connect()`
function. This will display a helpful message detailing which software package
was connected to R. It also returns a helpful object that we will use later.

```{r}
library(caliperR)
dk <- connect()
```

If you have multiple software applications installed, you can use `connect()` to
specify which you want to connect to.

```{r, eval=FALSE}
dk <- connect("Maptitude")
```

The only part of connecting that can be tricky is if you have multiple versions
of the same software installed (e.g. TransCAD 8 and TransCAD 9). By default,
`caliperR` selects the last one installed. To connect to a different version,
you must open that application before connecting. This can be done from R (but
manual works, too):

```{r, eval=FALSE}
system2(
  "C:\\Program Files\\TransCAD 9.0\\tcw.exe",
  wait = FALSE, invisible = FALSE
)
dk <- connect()
```

To disconnect (and close the software), use:

```{r, eval=FALSE}
disconnect()
```

## Macros and Functions

Caliper software comes with a scripting language called GISDK. Users write custom
macros in this language for various applications, but it also comes with a large
number of built-in functions.

### Running a macro

Running a GISDK macro that you or someone else has written in easy and looks
just like it would in GISDK.

```{r}
RunMacro("G30 Tutorial Folder")
```

### Running a function

To access the library of native GISDK functions, use the `dk` object returned by
`connect()`.

```{r}
folder <- RunMacro("G30 Tutorial Folder")
view <- dk$OpenTable("airports", "ffb", list(paste0(folder, "\\airports.bin")))
num_rows <- dk$GetRecordCount(view, NA)
paste0("The table '", view, "' has ", num_rows, " rows.")

# Equivalent in GISDK
# view = OpenTable("airports", "ffb", {folder + "airports.bin", null}, )
# num_rows = GetRecordCount(view, )
# ShowMessage("The table '" + view + "' has " + String(num_rows) + " rows.")
```

The code above introduces two other important differences between R and GISDK:

  * Use a `list` in R when you would use an `array` in GISDK.
  * Use `NA` in R when you would use `null` in GISDK.

To improve integration with R, select GISDK functions have been implemented
directly in the `caliperR` package:

  * `RunMacro()`
  * `SetAlternateInterface()`
  * `GetInterface()`
  * `CreateObject()`

These functions behave the same whether called directly or using the `dk`
object.

```{r, error=TRUE}
GetInterface()
dk$GetInterface()
```

## Custom Macros

In order to call one of your own GISDK functions from R, it must be compiled.
The `caliperR` package comes with the utility function `compile_gisdk()` to make
this easy and a toy script to test it on. Code compiled using Caliper software
also works.

```{r}
toy_script <- system.file("extdata", "gisdk", "testing", "gisdk.rsc", package = "caliperR")
ui_path <- compile_gisdk(toy_script, tempfile(fileext = ".dbd"))
```

The toy functions are simple, but illustrate some important points in the
examples below.

To reference a custom function that you have written and compiled, first use 
`SetAlternateInterface()` to point to your compiled code. The "first macro"
example in the compiled toy script simply returns a "Hello World!" message.

```{r}
SetAlternateInterface(ui_path)
RunMacro("first macro")
```

You can always check the current interface using

```{r}
GetInterface()
```

and set it back to the default interface with

```{r}
SetAlternateInterface()
GetInterface()
```

## Named Arrays

Many GISDK functions and macros take named arrays (aka options arrays) as input.
`caliperR` allows you to pass named lists as an equivalent.

```{r}
SetAlternateInterface(ui_path)

opts <- list()
opts$one <- 1
opts$two <- 2
RunMacro("parse opts array", opts)
```

## Views in R

Caliper software represents tabular data in "views". An earlier example
explained how to use `dk$OpenTable()` to open a view and get the number of rows
using GISDK functions. It can also be helpful to bring that tabular data into R
for additional processing. Two functions exist to help with this:

  * `df_from_view()`: convert a Caliper view into an R data frame
  * `df_to_view()`: convert a data frame into a Caliper view (or update an
    existing view)

In the code below, a data.frame is sent to a view. A new view is created, and
it's name is returned.
  
```{r}
SetAlternateInterface() # reset to default
df <- data.frame("one" = c(1, 2), "two" = c(3, 4), "three" = c(5, 6))
df
view <- df_to_view(df)
view
```
  
GISDK functions will now operate on that view as seen in previous examples.

```{r}
dk$GetRecordCount(view, NA)
```

The view can be brought back into a new data.frame.

```{r}
df2 <- df_from_view(view)
df2
```

Finally, the view can be updated with new data. To do so, we include the name
of the view we want to update in our call to `df_to_view()`.

```{r}
df2$one <- 10
view <- df_to_view(df2, view)
df_from_view(view)
```

## Logging and Message Boxes

The behavior of message boxes (errors or otherwise), depends on how Caliper
software is being run from R. Most often, the software is running in the
background (there is no icon on the task bar and no associated window). In this
case, messages are written out to a log file. In the code below, I first write
something to the log and then use `read_log()` to display it.

```{r, eval=FALSE}
dk$ShowMessage("This is a test.")
read_log()
```

If a Caliper program is already open when you connect to it, then it will have a
visible window and show up on the task bar. In this case, `ShowMessage` and
related functions will create a window with an OK button. R will wait until you
have closed that pop up before continuing. You can test the behavior by running
the code below. You may have to alt-tab to find the pop up window.

```{r, eval=FALSE}
dk$ShowArray(list(1, 2, 3))
```

## Advanced topics

For advanced topics, see the other package vignettes:

  * [Matrices](caliper-matrices.html)
  * [Objects](caliper-objects.html)
  * [Bin Files](bin-files.html)
